---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. Here's an example using the Lahman database. 

First, load the database, along with the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's look at an example. 

If I wanted to see **total homeruns** for the 1927 Yankees, I would have to extract data *FROM* the Batting table, and identify what specific data I want to limit my extraction to using the *WHERE* promt.

In this instance, I would use the following query:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)

```

## Greater Than, Less Than, or Equal To

We can also use equal to (=), greater than (>), or less than (<) symbols to narrow in our results. Lets look at a few examples. 

Say I wanted to find all instances where the Yankees hit **40 homeruns or more**:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)

```

To find all the instances in history where a player had **more than 40 homeruns** (HR), but **less than 40 strikeouts** (SO), I would enter the following query:

```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)

```

Say I wanted to look at a **specific time period**. For example, I want to find all instances of the Phillies hitting more than 30 homeruns **in the 1970s**:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<1980 and HR>30"
sqldf(query)

```

## ORDER BY 
#### Order the rows of your output

Let's look at how we can organize these statistics in a specific order. 

I want to find every instance of a player hitting more than 50 homeruns, but I want to *ORDER BY* having the player with **highest homerun total at the top**. To get the highest number at the top, we have to add *DESC* at the end of the *ORDER BY* prompt.

The query should look something like this:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)

```

Let's look at an example where we do not use the *DESC* prompt.

Say I wanted to find **all instances of a player striking out less than 10 times**. However, I want each player to have had **at least 400 at-bats (AB)**. I want to *ORDER BY* having the least amount of strike outs (SO) at the top.

```{r}
query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"

sqldf(query)

```

## Aggregation

Aggregation prompts allow us to find the sum *(sum)*, average *(avg)*, maximum *(max)*, or minimum *(min)* in a specific column. We do this by entering a query where we specify which column we want to apply the prompt to. For this to be effective, we have to use the *GROUP BY* prompt. This tells the program how we would like our information to be condensed.

Let's try applying this. 

Say we wanted to find all Babe Ruth's **total homeruns**. We would do this by writing in *sum(HR)*, followed by *GROUP BY* at the end of the query. 

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)

```

Another example of this uses a slightly different method of condensing information. In the following example, we will be using the prompt *HAVING*. This is essentially the same as the previous promt *WHERE*. However, it is used in instances where we are using aggregation prompts. 

Let me explain. We use *WHERE* when we want to specify a yearID, playerID, etc. (e.g. WHERE yearID=1980). *WHERE* always comes before *GROUP BY*. We use *HAVING* when we want to specify a sum(colname), avg(colname), max(colname), or min(colmame). Let's look at an example of this.

Let's find **all career homerun totals for all players**, but limit the output to **totals 600 or more**. Order by career homerun total.

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)

```

Here's another example. I want to find the players who have **averaged more than 30 homeruns per year** throughout their career. Order them by having the highest average at the top.

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query)

```

### INNER JOIN

To join two tables together, we would have to use the prompt *INNER JOIN*. This prompt allows us to extract data from multiple tables at once. However, it is important to specify in your query where the information you want should be extracted from. 

As an example, say we are using the Batting table and the Master table. We specify this by typing out the prompt like this: **Batting.colname**. To make sure that data isn't duplicated, we use the prompt *ON* and equate the colnames that overlap in each table to one another, like this: **ON Batting.colname=Master.colname**

Let's look at a few examples.

In the following query, we want to find:

**First and Last names, along with team, year, and homeruns**. We want the player to be **Babe Ruth**. Find all instances of players hitting **more than 50 homeruns**. 

```{r}
query<-"SELECT nameFIRST, nameLAST,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)

```

In another example, we want to find all instances of players hitting **more than 50 homeruns**. Include **first and last names, team, year, and homeruns**.

```{r}
query<-"SELECT nameFIRST, nameLAST,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50"

sqldf(query)

```

Let's look at a final example. Here we will be using the Batting and Teams tables. The Teams table has some overlapping information with Batting. Therefore, we need to indicate where we want to pull the information from. 

REMEMBER, when we use *ON*, we need to indicate what information is the same in both tables and eliminate duplicates.

We want to find **Babe Ruth's homeruns by season**, with playerID, team name, year, and homeruns. 

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR 
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)

```

### Joining Multiple Tables

Say we want to join the Batting table, Master table, and Teams tables all in one query. To be able to specify the data without there being any confusion, we have to utilize parenthesis () between each joining of tables.

For example, we want to see **Babe Ruth's first and last name, the year, homeruns and team name**.

```{r}
query<-"SELECT nameFIRST,nameLAST,name,Batting.yearID,Batting.HR
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)

```

Here's another example. Let's get the career **homerun totals for players**. Limit the display to just those that hit **at least 600HR**. Order by highest homerun totals at the top. Make sure we add the **first and last names, and HR total**.


```{r}
query<-"SELECT nameFIRST,nameLAST,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)

```

Here's a great example of what we talked about earlier with *WHERE* and *HAVING*. In this eample, we want to list **all players with the first name Bob** who have **averaged more than one millin dollars in salary** over their career. We want to include last names.

REMEMBER, *WHERE* before *GROUP BY* and *HAVING* (avg, sum, max, min) after.


```{r}
query<-"SELECT nameFIRST,nameLAST,avg(salary)
FROM Master INNER JOIN Salaries
ON Master.playerID=Salaries.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)

```

