---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. Here's an example using the Lahman database. First, load the database, along with the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

If I wanted to see **total homeruns** for the 1927 Yankees, I could write the following query:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)

```

We can also use greater than or less than symbols to narrow in our results. Lets look at a few examples. Say I wanted to find all instances where the Yankees hit **40 homeruns or more**:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)

```

To find all the instances in history where a player had **more than 40 homeruns**, but **less than 40 strikeouts**, I would enter the following query:

```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)

```

Say I wanted to look at a **specific time period**. For example, I want to find all instances of the Phillies hitting more than 30 homeruns **in the 1970s**:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<1980 and HR>30"
sqldf(query)

```

## ORDER BY 
#### Order the rows of your output

Let's look at how we can organize these stats in a specific order. I want to find every instance of a player hitting more than 50 homeruns, but I want to order them by having the player with **highest homerun total at the top**:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)

```

Find all instances of a player striking out less than 10 times. Make sure each player has had at least 400 at-bats (AB). Order by having least strikeouts at the top

```{r}
query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"

sqldf(query)

```

### Aggregatior

We want to find all Babe Ruth's total homeruns 

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)

```

Find all career homerun totals for all players, but limit the output to totals 600 or more. Order by career total.

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)

```

Find the players who have averaged more than 30 homeruns per year throughout their career. Order them by having the highest average at the top.

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query)

```

